\documentclass[11pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath, amssymb, amsthm}
\usepackage{booktabs}
\usepackage{multirow}

\newcommand{\Usingle}{\mathcal{U}^{\text{single}}}
\newcommand{\UsNO}{\mathcal{U}^{s}_{NO}}
\newcommand{\UsO}{\mathcal{U}^{s}_{O}}
\newcommand{\UD}{\mathcal{U}^{D}}
\newcommand{\UDA}{\mathcal{U}^{D}_{A}}
\newcommand{\Umax}{\mathcal{U}^{\max}}
\newcommand{\Ugreedy}{\mathcal{U}^{\text{greedy}}_{A}}

\title{Computational Results: Dynamic Augmented Pooled Testing}
\author{Vladimir, Hector, Francisco}
\date{March 2026}

\begin{document}
\maketitle

\section{Strategy Definitions}

We consider six testing strategies for a population instance $J = (p_1, \ldots, p_n, u_1, \ldots, u_n)$
with budget $B$ and maximum pool size $G$:

\begin{enumerate}
    \item $\Usingle(J, B, G)$: Optimal strategy testing single individuals.
    Each test contains exactly one individual. Select the $\min(B, n)$
    individuals with highest $u_i \cdot q_i$.

    \item $\UsNO(J, B, G)$: Optimal \textbf{static non-overlapping} pooled tests
    (classical binary $+/-$ results). Choose $B$ disjoint pools of size $\leq G$
    before seeing any results. Individual $i$ is cleared if their pool tests negative.

    \item $\UsO(J, B, G)$: Optimal \textbf{static overlapping} pooled tests
    (classical binary $+/-$ results). Choose $B$ pools of size $\leq G$
    (pools may share individuals). Individual $i$ is cleared if \emph{any}
    pool containing $i$ tests negative.

    \item $\UD(J, B, G)$: Optimal \textbf{dynamic} pooled tests with
    \textbf{classical} binary results ($+/-$). The strategy adapts: the pool
    at step $k$ depends on previous outcomes. Test result is
    $r \in \{-, +\}$ where $-$ means ``no infected in pool.''

    \item $\UDA(J, B, G)$: Optimal \textbf{dynamic augmented} pooled tests.
    Same as $\UD$ but the test returns the \emph{exact count}
    $r = |t \cap Z| \in \{0, 1, \ldots, |t|\}$ of infected individuals in the pool.
    Strictly more informative than classical binary results.

    \item $\Umax(J, B, G) = \sum_{i=1}^{n} u_i \cdot q_i$: Upper bound
    achieved only with infinite budget.
\end{enumerate}

\begin{theorem}[Inequality Chain]
For any population instance $J$, budget $B \geq 1$, and pool size $G \geq 1$:
\[
    \Usingle \;\leq\; \UsNO \;\leq\; \UsO \;\leq\; \UD \;\leq\; \UDA \;\leq\; \Umax
\]
\end{theorem}

Additionally, we define a \textbf{myopic augmented greedy} strategy $\Ugreedy$
that at each step selects the pool maximizing immediate expected utility gain
$\prod_{i \in t} q_i \cdot \sum_{i \in t} u_i$, then updates beliefs via
Bayesian posterior inference using the augmented test count. By construction,
$\Usingle \leq \Ugreedy \leq \UDA$.


\section{Bayesian Posterior Update}

A key component of the augmented greedy is the Bayesian update after observing
$r = |t \cap Z|$ for a tested pool $t$. For each individual $i \in t$, the
posterior infection probability is:
\[
    p'_i = \mathbb{P}(Z_i = 1 \mid r) = \frac{\mathbb{P}(r \mid Z_i = 1) \cdot p_i}
    {\mathbb{P}(r \mid Z_i = 1) \cdot p_i + \mathbb{P}(r \mid Z_i = 0) \cdot q_i}
\]
where, letting $S = t \setminus \{i\}$:
\begin{align*}
    \mathbb{P}(r \mid Z_i = 1) &= \mathbb{P}\!\left(\sum_{j \in S} Z_j = r - 1\right)
    \quad \text{(Poisson-Binomial over $S$)} \\
    \mathbb{P}(r \mid Z_i = 0) &= \mathbb{P}\!\left(\sum_{j \in S} Z_j = r\right)
\end{align*}
For $i \notin t$, the test provides no information: $p'_i = p_i$.

\noindent Special cases:
\begin{itemize}
    \item $r = 0$: All pool members are proven healthy ($p'_i = 0$ for all $i \in t$).
    \item $r = |t|$: All pool members are proven infected ($p'_i = 1$ for all $i \in t$).
\end{itemize}


\section{Computational Results}

All values computed via brute-force exact enumeration over all $2^n$ infection profiles.

\subsection{Full Inequality Chain}

\begin{table}[h]
\centering
\caption{Comparison of all strategies across four instances.}
\label{tab:inequality}
\begin{tabular}{@{}lcccccc@{}}
\toprule
 & $\Usingle$ & $\UsNO$ & $\UsO$ & $\UD$ & $\UDA$ & $\Umax$ \\
\midrule
Instance 1 & 11.84 & 18.81 & 19.01 & 19.01 & 19.07 & 22.19 \\
Instance 2 & 8.00  & 8.60  & 8.60  & 8.67  & 8.91  & 12.40 \\
Instance 3 & 7.30  & 8.56  & 8.91  & 8.91  & 8.98  & 9.70 \\
Instance 4 & 1.70  & 3.29  & 3.29  & 3.29  & 3.29  & 4.25 \\
\bottomrule
\end{tabular}
\vspace{0.5em}

\footnotesize
Instance 1: $n{=}5$, $B{=}2$, $G{=}3$, $p{=}(0.05, 0.10, 0.15, 0.20, 0.08)$, $u{=}(4, 6, 3, 5, 7)$. \\
Instance 2: $n{=}4$, $B{=}2$, $G{=}2$, $p{=}(0.30, 0.40, 0.35, 0.25)$, $u{=}(5, 3, 4, 6)$. \\
Instance 3: $n{=}3$, $B{=}2$, $G{=}2$, $p{=}(0.10, 0.20, 0.30)$, $u{=}(5, 3, 4)$. \\
Instance 4: $n{=}5$, $B{=}2$, $G{=}3$, $p{=}(0.15, \ldots, 0.15)$, $u{=}(1, \ldots, 1)$.
\end{table}

\subsection{Augmented Benefit}

The augmented benefit measures how much the exact count information helps
compared to classical binary test results in the dynamic setting:
\[
    \text{Augmented Benefit} = \frac{\UDA - \UD}{\UD} \times 100\%
\]

\begin{table}[h]
\centering
\caption{Augmented benefit over classical dynamic testing.}
\label{tab:augmented-benefit}
\begin{tabular}{@{}lcccl@{}}
\toprule
 & $\UD$ & $\UDA$ & Benefit & Remark \\
\midrule
Instance 1 & 19.01 & 19.07 & $+0.28\%$ & Low infection rates \\
Instance 2 & 8.67  & 8.91  & $+2.74\%$ & High infection rates \\
Instance 3 & 8.91  & 8.98  & $+0.81\%$ & Mixed rates \\
Instance 4 & 3.29  & 3.29  & $+0.00\%$ & Uniform population \\
\bottomrule
\end{tabular}
\end{table}

Key observation: the augmented benefit is \textbf{largest when infection rates are high}
(Instance~2: $+2.74\%$). This aligns with the paper's intuition that augmented
information tests are more informative precisely when infection rates increase,
unlike classical protocols like Dorfman's which degrade.

For uniform populations (Instance~4), there is no benefit: symmetry means the
count provides no additional information about \emph{who} is infected.


\subsection{Greedy Performance}

\begin{table}[h]
\centering
\caption{Myopic augmented greedy vs.\ optimal augmented dynamic.}
\label{tab:greedy}
\begin{tabular}{@{}lcccl@{}}
\toprule
 & $\Ugreedy$ & $\UDA$ & Gap & \% of optimal gap captured \\
\midrule
Instance 1 & 19.053 & 19.066 & 0.013 & 99.8\% \\
Instance 2 & 8.876  & 8.909  & 0.033 & 96.3\% \\
Instance 3 & 8.966  & 8.982  & 0.016 & 99.1\% \\
Instance 4 & 3.287  & 3.287  & 0.000 & 100.0\% \\
\bottomrule
\end{tabular}
\vspace{0.5em}

\footnotesize
``\% of optimal gap captured'' $= (\Ugreedy - \Usingle) / (\UDA - \Usingle) \times 100\%$.
\end{table}

The myopic greedy with augmented Bayesian updates captures over 96\% of the
gap between individual testing and optimal in all instances. This suggests
that simple greedy policies with augmented posterior updates may be
near-optimal in practice.


\subsection{Simulation Example}

Consider Instance~1 with infection profile $Z = \{3\}$ (only individual~3
is infected).

\paragraph{Optimal strategy ($\UDA$).}
\begin{enumerate}
    \item Test pool $\{2,3,4\}$, result $r = 1$.
    \item Bayesian posteriors: $p'_2 = 0.34$, $p'_3 = 0.49$, $p'_4 = 0.17$.
    \item Test pool $\{0,1,4\}$, result $r = 0 \to$ \textsc{cleared}.
    \item Cleared: $\{0, 1, 4\}$, utility $= 17.0$.
\end{enumerate}

\paragraph{Myopic greedy.}
\begin{enumerate}
    \item Test pool $\{0,1,4\}$ (highest immediate score), result $r = 0 \to$ \textsc{cleared}.
    \item Test pool $\{2,3\}$, result $r = 1$.
    \item Cleared: $\{0, 1, 4\}$, utility $= 17.0$.
\end{enumerate}

Both achieve the same utility on this profile, but the optimal strategy
\emph{tests the riskier pool first} to gain information, while the greedy
\emph{tests the safest pool first} for immediate gain. The optimal strategy's
advantage appears in expectation over all possible infection profiles.


\section{Robustness Under High Infection Rates}

A recurring concern in classical pooled testing is that protocols
degrade as infection rates increase. For example, Dorfman's two-stage
procedure performs well when the prevalence is low (e.g., $p < 0.1$)
but rapidly loses efficiency as $p$ grows, because most pools test
positive and require individual retesting.

In the augmented setting, however, even a positive pool provides
quantitative information: the exact count $r = |t \cap Z|$ narrows
the set of consistent infection profiles. This leads to the following
hypothesis, originally stated in the overview:

\begin{quote}
\emph{In the augmented pooled testing regime, the benefit over
classical dynamic testing may persist --- or even increase --- as
infection rates grow.}
\end{quote}

\subsection{Experimental Design}

We define four infection rate \emph{regimes} by restricting the prior
infection probabilities $p_i$ to specific intervals:

\begin{center}
\begin{tabular}{@{}lcc@{}}
\toprule
\textbf{Regime} & \textbf{Range of $p_i$} & \textbf{Avg.\ $p$} \\
\midrule
Low        & $[0.05, 0.15]$ & $\approx 0.10$ \\
Medium     & $[0.15, 0.35]$ & $\approx 0.25$ \\
High       & $[0.35, 0.55]$ & $\approx 0.45$ \\
Very High  & $[0.55, 0.80]$ & $\approx 0.68$ \\
\bottomrule
\end{tabular}
\end{center}

For each regime we generate 200 random instances with $n = 5$,
$B \in \{2, 3\}$, $G = 3$, and compute both the optimal classical
dynamic utility $\UD$ and the optimal augmented dynamic utility $\UDA$
via exact brute-force enumeration. We measure the \emph{augmented
benefit} as defined in Section~3.2.

\subsection{Results}

\begin{table}[h]
\centering
\caption{Augmented benefit by infection rate regime ($n{=}5$, $G{=}3$, 200 instances per regime).}
\label{tab:high-infection}
\begin{tabular}{@{}lcccc@{}}
\toprule
\textbf{Regime} & \textbf{Mean Benefit} & \textbf{Max Benefit} & \textbf{Std} & \textbf{Instances with Benefit $> 0$} \\
\midrule
Low        & $+0.26\%$ & $+3.8\%$  & $0.5\%$ & $41\%$ \\
Medium     & $+0.92\%$ & $+5.1\%$  & $1.2\%$ & $63\%$ \\
High       & $+0.67\%$ & $+4.2\%$  & $0.9\%$ & $58\%$ \\
Very High  & $+0.00\%$ & $+0.0\%$  & $0.0\%$ & $0\%$  \\
\bottomrule
\end{tabular}
\end{table}

\subsection{Discussion}

Three observations emerge from Table~\ref{tab:high-infection}:

\begin{enumerate}
    \item \textbf{The augmented benefit peaks at medium infection rates.}
    The mean benefit is highest in the medium regime ($+0.92\%$), where
    the count information $r$ is most discriminative: pools frequently
    test positive with $r \in \{1, 2\}$, allowing the planner to
    distinguish between infection profiles that classical binary tests
    conflate into a single ``$+$'' result.

    \item \textbf{At very high rates, both models collapse.}
    When $p_i > 0.55$ for all agents, virtually every pool returns
    $r \approx |t|$, providing little diagnostic value. In this regime,
    the augmented count offers no advantage because the space of
    consistent profiles is already tightly constrained --- or
    equivalently, almost no individual can be proven healthy under
    either model.

    \item \textbf{The augmented model does not degrade like classical protocols.}
    Unlike Dorfman's scheme --- where efficiency drops sharply past a
    critical prevalence --- the augmented benefit remains positive
    across the low, medium, and high regimes. The degradation only
    occurs in the very high regime, where the problem itself becomes
    fundamentally harder (few individuals can be cleared regardless
    of the testing model).
\end{enumerate}

These results provide empirical evidence supporting the hypothesis
that augmented pooled testing is more robust to increases in infection
rates than classical pooled testing. The key mechanism is that the
count $r$ provides a finer partition of the outcome space: for a pool
of size $G$, augmented tests distinguish $G + 1$ outcomes versus the
$2$ outcomes of classical tests, and this distinction is most valuable
at intermediate prevalence levels where the outcomes are neither
trivially all-negative nor all-positive.


\end{document}
